A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-05-09 graph_syntax.tex
removing old Fig 1.

  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) 
  and one edge type are included \cite{alexandrov2004}. We have chosen to adopt 
  the terminology ``model'' node instead of ``function'' node to be consistent 
  with modern MDAO framework terminology. The present work adds one additional 
  node type and two additional edge types to the syntax to allow descriptive
  graphs for all three phases of the design problem formulation process indicated in Figs.~\ref{f:tree} and \ref{f:hourglass}.
\begin{comment}
In addition, we introduce the 
  concept of fixed vs. free connection edges to represent the
  static relationship between variable and model nodes from a single analysis, 
  as opposed to the flexible relationship between variables from different analyses. 
\end{comment}

@@

  All global variable nodes have more than one outgoing edge. Conversely all local 
  variable nodes have only a single outgoing edge. Fig.~\ref{f:sellar graph} contains 
  three global variables that are not present in the problem formulation given by 
  Eq.~\ref{eqn:simple_fpf}: $x_1^*$, $z_1^*$, and $z_2^*$. The presence of the 
  new nodes is required in order to represent the equality of the given variables that
  was implied by their shared names in Eq.~\ref{eqn:simple_fpf}.

  The variable $x_1$ is usually refered to as a local variable, which is 
  contradicted by the given graph \cite{sellar1996}. This discrepancy 
  arises from the explicit treatment of objectives and constraints as separate 
  expression blocks. This variable is could be considered a local 
  variable because it directly affects only analysis block $A_1$. However, by expanding
  $F$ into an expression block with its own inputs, $F$ maintains its own local 
  copy of the $x_1$ variable. This formulation necessitates the creation of the global 
  node $x_1^*$ in the graph with two outgoing edges to link the two $x_1$ inputs nodes 
  in the respective blocks.

@@

  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
  definition; however, one or more solvers will be necessary to build a valid PSG from a given FPG. 
  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select
  node in the cycle as a starting point and proceed around the
  loop until the starting point is reached again. For the Sellar Problem, selecting 
  $A_1$ as the starting point would yield a problem as given in 
  Eq.~\ref{eqn:simple}, whereas selecting $A_2$ would yield the problem as given in 
  Eq.~\ref{eqn:simple2}.

  Larger problems can contain more complex cycles in their FPG, indicating more 
  complex coupling between analyses. For example, a cycle can involve more than 
   two analysis blocks. Multiple independent cycles could also exist, indicating 
  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. In the present work, Sec.~\ref{s:example problem}
  demonstrates how building an FPG from an MCG provides an opportunity to 
  identify and potentailly reduce the number of cycles in a graph.

>>>>

  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) 
  and one edge type are included \cite{alexandrov2004}. We have chosen to adopt 
  the terminology ``model'' node instead of ``function'' node to be consistent 
  with modern MDAO framework terminology. The present work adds one additional 
  node type and two additional edge types to the syntax to allow descriptive
  graphs for all three phases of the design problem formulation process indicated in Figs.~\ref{f:tree} and \ref{f:hourglass}.

@@

  All global variable nodes have more than one outgoing edge. Conversely all local 
  variable nodes have only a single outgoing edge. Fig.~\ref{f:sellar types} contains 
  three global variables that are not present in the problem formulation given by 
  Eq.~\ref{eqn:simple_fpf}: $x_1^*$, $z_1^*$, and $z_2^*$. The presence of the 
  new nodes is required in order to represent the equality of the given variables that
  was implied by their shared names in Eq.~\ref{eqn:simple_fpf}.

  The variable $x_1$ is usually refered to as a local variable, which is 
  contradicted by the given graph \cite{Sellar1996}. This discrepancy 
  arises from the explicit treatment of objectives and constraints as separate 
  expression blocks. This variable is could be considered a local 
  variable because it directly affects only analysis block $A_1$. However, by expanding
  $F$ into an expression block with its own inputs, $F$ maintains its own local 
  copy of the $x_1$ variable. This formulation necessitates the creation of the global 
  node $x_1^*$ in the graph with two outgoing edges to link the two $x_1$ inputs nodes 
  in the respective blocks.

@@

  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
  definition; however, one or more solvers will be necessary to build a valid PSG from a given FPG. 
  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select
  node in the cycle as a starting point and proceed around the
  loop until the starting point is reached again. For the Sellar Problem, selecting 
  $D_1$ as the starting point would yield a problem as given in 
  Eq.~\ref{eqn:simple}.
  Larger problems can contain more complex cycles in their FPG, indicating more 
  complex coupling between analyses. For example, a cycle can involve more than 
   two analysis blocks. Multiple independent cycles could also exist, indicating 
  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. In the present work, Sec.~\ref{s:example problem}
  demonstrates how building an FPG from an MCG provides an opportunity to 
  identify and potentailly reduce the number of cycles in a graph.
